Parsing Postfix log files

### 2009-02-12 doc/logparser-postfix-daemons.tex
Fixed acronyms so they work properly, get expanded when appropriate, and add the     correct entries to the glossary and acronym appendices. Correct capitalisation of most section and subsection headings. Corrections made to background chapter; more remain to be made.

\newglossaryentry{lmtp}{name={lmtp},type={postfix},description={
    Delivery of mail over \gls{LMTP} is performed by the lmtp component.
    The definitive documentation is
    \url{http://www.postfix.org/lmtp.8.html}.
}}

@@

\newglossaryentry{smtp}{name={smtp},type={postfix},description={
    Delivery of mail over \gls{SMTP} is performed by the smtp component.
    The definitive documentation is
    \url{http://www.postfix.org/smtp.8.html}.
}}

\newglossaryentry{smtpd}{name={smtpd},type={postfix},description={
    Smtpd is the Postfix program that accepts mail via \gls{SMTP}, and
    implements all the anti-spam restrictions Postfix provides.  The
    definitive documentation is \url{http://www.postfix.org/smtpd.8.html}.
}}

>>>>

\newglossaryentry{lmtp}{name={lmtp},type={postfix},description={
    Delivery of mail over \acronym{LMTP} is performed by the lmtp
    component.  The definitive documentation is
    \url{http://www.postfix.org/lmtp.8.html}.
}}

@@

\newglossaryentry{smtp}{name={smtp},type={postfix},description={
    Delivery of mail over \acronym{SMTP} is performed by the smtp
    component.  The definitive documentation is
    \url{http://www.postfix.org/smtp.8.html}.
}}

\newglossaryentry{smtpd}{name={smtpd},type={postfix},description={
    Smtpd is the Postfix program that accepts mail via \acronym{SMTP}, and
    implements all the anti-spam restrictions Postfix provides.  The
    definitive documentation is \url{http://www.postfix.org/smtpd.8.html}.
}}

### 2009-02-12 doc/logparser-results.tex
Dealt with some XXX items.
Fixed acronyms so they work properly, get expanded when appropriate, and add the     correct entries to the glossary and acronym appendices. Correct capitalisation of most section and subsection headings. Corrections made to background chapter; more remain to be made.

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that as expected there are far more
mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \gls{SMTP} only; in particular the mail loops noticeable
in other graphs do not contribute to these figures.

@@

\subsection{Explaining the peaks in log file size}

@@

\subsection{Rule ordering for efficiency}

@@

\subsection{Scalability as the number of rules rises}

@@

\newpage{} % XXX
\subsection{Caching compiled regexs}

@@

\subsection{Log lines covered}

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \gls{DNSBL} messages change over time, major releases of Postfix
change warning messages (usually adding more information), etc.,\ so over
time the log lines drift and change.  \Graphref{rule hits graph} shows the
number of hits for each rule over all \numberOFlogFILES{} log files.  The
number of hits per rule is quite unevenly spread, resembling a Power Law
distribution --- it is obvious that a small number of rules match the vast
majority of the log lines, and more than half the rules match fewer than
100 times.

@@

\subsection{Mails covered}

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails that were not handled properly (many of the
complications discussed in \sectionref{complications} were
detected in this way).  The parser issues warnings when it detects any
errors, some of which may alert the user to a problem, e.g.\ when a queueid
is reused before the previous mail is fully dealt with, when a queueid or
\gls{pid} is not found in the state tables,\footnote{There will often be
warnings about a missing queueid or \gls{pid} in the first few hundred or
thousand log lines because the earlier log lines for those connections or
mails are in the previous log file; loading the saved state from the
previous log file will solve this problem.} or when there are problems
tracking a child mail (see \sectionref{tracking re-injected mail}).  There
should be few or no warnings when parsing, and when finished parsing the
state table should only contain entries for mails that had yet to be
delivered when the log files ended, or were accepted before the log files
began.

>>>>

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that as expected there are far more
mails received on weekdays than at weekends.   Median log file size is
50MB, containing 285,000 log lines --- large scale mail servers would have
much larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \acronym{SMTP} only; in particular the mail loops
noticeable in other graphs do not contribute to these figures.

@@

\subsection{Explaining the Peaks in Log File Size}

@@

\subsection{Rule Ordering For Efficiency}

@@

\subsection{Scalability as the Number of Rules Rises}

@@

\subsection{Caching Compiled Regexs}

@@

\subsection{Log Lines Covered}

@@

Full coverage of log lines is easy to achieve yet hard to maintain.  It is
easy to achieve full coverage for a limited set of log files (at the time
of writing the parser is tested with \numberOFrules{} rules, fully parsing
\numberOFlogFILESall{} contiguous log files from Postfix 2.2 through to
Postfix 2.5), and new rules are easy to add.  Maintaining full coverage is
hard because other servers have different restrictions with custom
messages, \acronym{DNSBL} messages change over time, major releases of
Postfix change warning messages (usually adding more information), etc.,\
so over time the log lines drift and change.  \Graphref{rule hits graph}
shows the number of hits for each rule over all \numberOFlogFILES{} log
files.  The number of hits per rule is quite unevenly spread, resembling a
Power Law distribution --- it is obvious that a small number of rules match
the vast majority of the log lines, and more than half the rules match
fewer than 100 times.

@@

\subsection{Mails Covered}

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails that were not handled properly (many of the
complications discussed in \sectionref{complications} were detected in this
way).  The parser issues warnings when it detects any errors, some of which
may alert the user to a problem, e.g.\ when a queueid is reused before the
previous mail is fully dealt with, when a queueid or \acronym{pid} is not
found in the state tables,\footnote{There will often be warnings about a
missing queueid or \acronym{pid} in the first few hundred or thousand log
lines because the earlier log lines for those connections or mails are in
the previous log file; loading the saved state from the previous log file
will solve this problem.} or when there are problems tracking a child mail
(see \sectionref{tracking re-injected mail}).  There should be few or no
warnings when parsing, and when finished parsing the state table should
only contain entries for mails that had yet to be delivered when the log
files ended, or were accepted before the log files began.
